A Way Forward in Parallelising Dynamic Languages

### 2014-06-18 talk/icooolps2014/position-paper.tex
Two minor fixes (too late?  never mind)

\subsection{Why is there a GIL?}
The GIL is a very simple synchronisation mechanism for supporting
multithreading in an interpreter. The basic guarantee is that the GIL
may only be released in between bytecode instructions\footnote{This
also applies to Abstract Syntax Tree (AST) interpreters, where the GIL
may only be released between interpreting two AST nodes.}. The interpreter
can thus rely on complete isolation and atomicity for the
instructions' execution. Also, accesses to data structures like
dictionaries and lists happen atomically and do not need additional
protection from data races when shared between threads.

@@

\begin{description}
\item[Performance:] How much does the approach impact performance on a single
  and how much on multiple threads? Can it make use of parallelism?
\item[Existing applications:] How big are the changes required to
  integrate with and parallelise existing applications?
\item[Better synchronisation:] Does the approach enable better,
  parallelisable synchronisation mechanisms for applications
  (e.g.\ atomic blocks)?  Many synchronisation mechanisms can be built on
  top of all solutions (e.g.\  message passing, monitors). We look for mechanisms
  that are directly enabled by the contending approaches.
\item[Implementation:] How difficult is it to implement the approach
  in the interpreter?
\item[External libraries:] Does the approach allow for easy
  integration of external libraries?
\end{description}

>>>>

\subsection{Why is there a GIL?}
The GIL is a very simple synchronisation mechanism for supporting
multithreading in an interpreter. The basic guarantee is that the GIL
may only be released in between bytecode instructions\footnote{This
also applies to Abstract Syntax Tree (AST) interpreters, where the GIL
may only be released between interpreting two AST nodes.  We talk about
``bytecode instructions'' in a general way as a basic step in the
interpreter.}.  The interpreter
can thus rely on complete isolation and atomicity for the
instructions' execution. Also, accesses to data structures like
dictionaries and lists happen atomically and do not need additional
protection from data races when shared between threads.

@@

\begin{description}
\item[Performance:] How much does the approach impact performance
  on a single thread
  and how much on multiple threads? Can it make use of parallelism?
\item[Existing applications:] How big are the changes required to
  integrate with and parallelise existing applications?
\item[Better synchronisation:] Does the approach enable better,
  parallelisable synchronisation mechanisms for applications
  (e.g.\ atomic blocks)?  Many synchronisation mechanisms can be built on
  top of all solutions (e.g.\  message passing, monitors). We look for mechanisms
  that are directly enabled by the contending approaches.
\item[Implementation:] How difficult is it to implement the approach
  in the interpreter?
\item[External libraries:] Does the approach allow for easy
  integration of external libraries?
\end{description}
